Blame and Coercion: Together Again for the Second Time

### 2019-10-15 paper.tex
revise intro

Around 2006, many work on integrating dynamic typing and static typing emerge, 
notably gradual types \cite{siek2006gradual} and hybrid types 
\cite{flanagan2006hybrid}. 
Later \citet{wadler2009well} introduces Blame Calculus, an intermediate 
language for gradual types and hybrid types. 
Blame Calculus throws a blame label when a runtime type-check fails.
These labels can include useful information on the failed cast, for example, 
its location in the source code.

Implementing gradually typed languages naively on top of blame calculus 
suffices sever space leak.
For example, translation to Blame Calculus can wrap some expressions with a 
cast, including those at tail position. Thus, at runtime, mutually 
tail-recursive functions in source language can accumulate unbounded space. 
There are more examples in \citet{herman2010space}.

In 2007, \citet{herman2010space} proposes a solution to this problem by 
translating Blame Calculus to a variant of \citet{henglein1994dynamic}'s 
Coercion Calculus. Their key idea is to represent casts with coercions, 
which can be composed and normalized. Every coercion looks like a list of 
trees. For normal coercion, the length of the list is bounded by a small 
constant. And the depths of trees do not grow in composition and 
normalization. These two facts implies that coercion is a space-efficient cast 
representation. This Coercion Calculus, however, does not support blame 
tracking.

@@

\citet{siek2009exploring} incorporate blame tracking into 
\citet{herman2010space}'s Coercion Calculus by 
decorating coercions with labels. They also propose that there are four blame 
strategies for their Coercion Calculi:
$ \{Lazy, Eager\} \times \{D, UD\} $. 
Eager strategies detect more potential type errors, 
but also blame more programs.
$ D $ and $ UD $ assign blame labels differently.
They prove partially that the \lazyUD{} Coercion Calculus is correct by showing 
that it simulates the Blame Calculus in \cite{wadler2009well}. It is unknown, 
however, whether the simulation also works in the other direction.

\citet{siek2010threesomes} proposes another approach to combine blame 
tracking and space efficiency. Their solution is based on Threesome Calculus. 
Their key idea is to represent casts with threesomes, a novel cast 
representation.
%Threesomes can also be composed. Unlike coercions, there is no 
%separate normalization because every threesome is in normal form. 
They show that their threesome calculus bi-simulates the \lazyUD{} Coercion 
Calculus in \citet{siek2009exploring} and the Blame Calculus in 
\citet{wadler2009well}. They also prove that threesome is space-efficient as 
well.

\citet{siek2012interpretations} present the first \lazyD{} Blame Calculus and 
revised lazy Coercion Calculi.
They simplify the Coercion Calculi by only 
working with normal coercions (i.e. coercions in normal form). 
%Fortunately, coercions are trivially normal when initially constructed, and 
%their empowered compose function is conjectured to produces normal coercions 
%as well. 
They conjecture that their Coercion Calculi bi-simulate the corresponding Blame 
Calculi.

Following \citet{siek2010threesomes}, \citet{garcia2013calculating} discovers
Threesome Calculi for other blame strategies. After studying the relation 
between coercion and threesome,
he claims that normal coercion with labels is easy to understand but hard to 
implement and hard to reason about, and that 
threesome with labels, however, is easy to implement and easy to reason about 
but hard to understand. 
His claim is partially affirmed by the group of people who develop threesome 
(\citet{siek2015blame}).
His threesome calculi are equivalent to the Coercion Calculi in 
\citet{siek2009exploring} because the former ones are derived from the latter 
ones.

\citet{siek2015blame} revisit the coercion-based approach. They revise the 
\lazyUD{} Coercion Calculus in \citet{siek2012interpretations} and prove that 
their revised coercion calculus bi-simulates the \lazyUD{} Blame Calculus.

Last year, \citet{kuhlenschmidt2018efficient} present Grift, a compiler for a 
gradually typed language of the same name. 
This implementation is based on the \lazyD{} Coercion Calculus in 
\citet{siek2012interpretations}.
Their result suggests that normal coercion is easy enough to implement in a 
compiler.

@@

So far, it seems that the coercion-based approach is the best way to combine 
space-efficiency and blame-tracking: normal coercion is claimed easy to 
understand \cite{garcia2013calculating}\cite{siek2015blame} and is shown 
easy enough to implement in a compiler \cite{kuhlenschmidt2018efficient}.
Normal coercion is still not satisfactory, however, in at least three aspects.
Firstly, as we mentioned above, it is hard to reason about normal coercion. The 
major difficulty is from its non-structurally recursive composition.
For instance, many developers of Grift report it is tricky to convince
their proof assistants that the composition terminates. 
Secondly, the definition of normal coercion lies unnecessarily on top 
of coercion: all coercions are normal when they are initially constructed, and 
\citet{siek2012interpretations} have shown that there exist functions that 
compose and produce normal coercions.

Perhaps unsurprisingly, threesome has a structurally recursive compose and a 
self-standing definition. Together with \citet{garcia2013calculating}'s 
work, they suggest that there might be a cast representation whose definition 
is self-standing, and whose composition is structurally recursive. 
Super-coercion introduced in \citet{garcia2013calculating} could be a promising 
candidate. However, its definition is a bit complicated. It uses 10 
constructors to deal with an elementary type system with only base types and 
function types. And four constructors are directly related to function types. 
Thus, super-coercion might not scale very well to more sophisticated type 
systems.

We present hyper-coercion, a cast representation whose composition is 
structurally recursive, and whose definition is self-standing. What's more, it 
should be at 
least as easy to implement and understand as normal coercions, because there is 
a clear connection between them, as we will show in Section 
\ref{sec:hyper-coercion}.

Our hyper-coercion considers sum types and product types, which are not 
considered in all proofs above. Adding each of them requires us to add only one 
new constructor to a component of hyper-coercion. This suggests that 
hyper-coercion might scale better than super-coercion.

>>>>

Around 2006, many work on integrating dynamic typing and static typing emerge, 
notably gradual types \cite{siek2006gradual} and hybrid types 
\cite{flanagan2006hybrid}. 
Later \citet{wadler2009well} introduces Blame Calculus, an intermediate 
language for gradual types and hybrid types. 
Blame Calculus throws a blame label when a runtime type-check fails.
Labels provide useful debugging information, therefore are necessary for 
practical use of gradually typed languages.

Implementing gradually typed languages naively on top of blame calculus, 
however, suffices sever space leak.
For example, translation to Blame Calculus can wrap some expressions with code 
that performs runtime type-checking (casts), including those expressions at 
tail position. Thus, mutually tail-recursive functions in source language can 
consume unbounded space at runtime. There are more examples of space leaking in 
\citet{herman2010space}.

In 2007, \citet{herman2010space} proposes a solution to this problem by 
representing casts with coercion \cite{henglein1994dynamic}, 
which can be composed and normalized. Every coercion looks like a list of 
trees. For normal coercion, the length of the list is bounded by a small 
constant. And the depths of trees do not grow during composition and
normalization. These two facts implies that coercion is a space-efficient cast 
representation. This coercion, however, does not support blame 
tracking.

@@

\citet{siek2009exploring} decorate \citet{herman2010space}'s coercions with 
labels to restore 
blame-tracking. They also propose that there are four blame 
strategies for their coercion:
$ \{Lazy, Eager\} \times \{D, UD\} $. 
Eager strategies detect more potential type errors, 
but blame more programs.
$ D $ and $ UD $ assign blame labels differently.
They prove partially that their \lazyUD{} coercion is correct by showing 
that it simulates the Blame Calculus in \cite{wadler2009well}. It is unknown, 
however, whether the simulation also works in the other direction.

\citet{siek2010threesomes} proposes another approach to combine blame 
tracking and space efficiency. Their solution represents casts with threesomes, 
a novel cast representation.
They show that their calculus bi-simulates the \lazyUD{} coercion 
calculus in \citet{siek2009exploring} and the Blame Calculus in 
\citet{wadler2009well}. They also prove that threesome is space-efficient as 
well.

\citet{siek2012interpretations} present the first \lazyD{} Blame Calculus and 
revised coercion calculi.
They simplify the coercion calculi by only 
working with normal coercions (i.e. coercions in normal form). 
They conjecture that their revised coercion calculi bi-simulate the 
corresponding Blame Calculi.

Following \citet{siek2010threesomes}, \citet{garcia2013calculating} develops
threesome for other blame strategies. After studying in depth the relation 
between coercion and threesome,
he claims that normal coercion with labels is easy to understand but hard to 
implement and hard to reason about, and that 
threesome with labels, however, is easy to implement and easy to reason about 
but hard to understand. 
His claim is mostly affirmed by the group of people who develop threesome 
(\citet{siek2015blame}).
His threesome calculi are equivalent to the coercion calculi in 
\citet{siek2009exploring} because the former ones are derived from the latter 
ones.

\citet{siek2015blame} revisit the coercion-based approach. They revise the 
\lazyUD{} coercion in \citet{siek2012interpretations} and prove that 
their revised coercion calculus bi-simulates the \lazyUD{} Blame Calculus.

Last year, \citet{kuhlenschmidt2018efficient} present Grift, a compiler for a 
gradually typed language of the same name. 
This implementation is based on the \lazyD{} coercion in 
\citet{siek2012interpretations}.
Their result suggests that normal coercion is easy enough to implement in a 
compiler.

@@

So far, it seems that normal coercion is the best way to combine 
space-efficiency and blame-tracking: normal coercion is claimed easy to 
understand \cite{garcia2013calculating}\cite{siek2015blame} and is shown 
easy enough to implement in a compiler \cite{kuhlenschmidt2018efficient}.
Normal coercion is still not satisfactory, however, in at least two aspects.
Firstly, as we mentioned above, it is hard to reason about normal coercion. The 
major difficulty is from its non-structurally recursive composition.
For instance, many developers of Grift report it is tricky to convince
their proof assistants that the composition terminates. 
Secondly, the definition of normal coercion lies unnecessarily on top 
of coercion: all coercions are normal when they are initially constructed, and 
\citet{siek2012interpretations} have shown an function that 
composes and produces normal coercions.

Perhaps unsurprisingly, threesome has a structurally recursive compose and a 
self-standing definition. Together with \citet{garcia2013calculating}'s 
work, they suggest that there might be a cast representation whose definition 
is self-standing, and whose composition is structurally recursive. 
Super-coercion introduced in \citet{garcia2013calculating} could be a promising 
candidate. It is believed even more complicated than threesome 
\cite{siek2015blame}. In fact, it uses 10 constructors to deal with an 
elementary type system with only base types and 
function types. And four constructors are directly related to function types. 
Thus, super-coercion might not scale very well to more sophisticated type 
systems.

We present hyper-coercion, a cast representation whose composition is 
structurally recursive, and whose definition is self-standing. What's more, it 
should be at 
least as easy to implement and understand as normal coercions, because there is 
a clear connection between them, as we will show in Section 
\ref{sec:hyper-coercion}.
Our hyper-coercion considers sum types and product types, which are not 
considered in all proofs above. Adding each of them requires us to add only one 
new constructor to a component of hyper-coercion. This suggests that 
hyper-coercion might scale better than super-coercion.
